---
title: "Data Visualzation with ggplot2 exercises"
author: "Fechar Ourotcha"
date: "3/27/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
data("mpg")
```
# EXERCISE 1

## Question 1) Run ggplot(data = mpg). What do you see? 
```{r, results='hide'}
ggplot(data = mpg)
```

When I run the code, I get a blank graph.

## Question 2) How many rows are in mtcars? How many columns? 
```{r results='hide'}
dim(mpg)
```
There are 234 rows and 11 columns in this dataset

## Question 3) What does the drv variable describe? Read the help for ?mpg ## to find out. 
```{r, results='hide'}
?mpg
```
the drv variable is a character class column; with information on whether the car model is either f, r, or 4 (f = front-wheel drive, r = rear wheel drive, 4 = 4w)

## Question 4) Make a scatterplot of hwy versus cyl.

```{r}
attach(mpg)
ggplot(data = mpg)+
  geom_point(mapping = aes( x = cyl, y = hwy ), color = ("green"))
```

# EXERCISE 2

## Question 1) What’s gone wrong with this code? Why are the points not  ## blue? 
```{r, results='hide'}
ggplot(data = mpg) + 
geom_point( mapping = aes(x = displ, y = hwy, color = "blue")  )


```
The color is not blue because the color argument is inside the aes, which is making r think "blue" is just a name. To change this bring the color argument outside of aes.

## Question 2) Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset.) How can you see this information when you run mpg? 
```{r, results='hide'}
glimpse(mpg)
```
Pretty much all the variables are categorical except for engine displacement, highway miles per gallon.

## Question 3) Map a continuous variable to color, size, and shape. How do ## these aesthetics behave differently for categorical versus continuous  ## variables? 


## Question 4) What happens if you map the same variable to multiple      ## aesthetics?

## Question 5) What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point.)
```{r}
?geom_point
```
the stroke aesthetic modifies the width of the border


# EXERCISE 3

## Quetsion 1) What happens if you facet on a continuous variable? 

when you facet on a continuous variable, you basically breack the Variable into different categories that ## allows you to get a more detail insight about that variable.

## Question 2) What do the empty cells in a plot with facet_grid(drv ~    ## cyl) mean? How do they relate to this plot? 

that means there are no value for the category. In the plot we see that there are no cars with 5 cylinders ## and 4 wheel drive. Also no cars with rear wheel drive and 4 or 5 cylinder.



## Question 3) What plots does the following code make? What does . do? 
```{r, results='hide'}
ggplot(data = mpg) +  geom_point(mapping = aes(x = displ, y = hwy)) +  facet_grid(drv ~ .)

ggplot(data = mpg) +  geom_point(mapping = aes(x = displ, y = hwy)) +  facet_grid(. ~ cyl)
```


The foolowing plots create scatter plots. the dot is just a place holder 


## Question 4) Take the first faceted plot in this section: 
```{r, results='hide'}
ggplot(data = mpg) +  geom_point(mapping = aes(x = displ, y = hwy)) +  facet_wrap(~ class, nrow = 2)
```

## What are the advantages to using faceting instead of the color aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

faceting allows you to see the different categories within a variable independently 

## Question 5) Read ?facet_wrap. 
## What does nrow do? 
## What does ncol do? 
## What other options control the layout of the individual panels? 
## Why doesn’t facet_grid() have nrow and ncol variables?

## Question 6). When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?

I guess because it looks better.

# EXERCISE 4

## Question 1) What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

line geom, histgram geom, boxplot geom, and an area geom

## Question 2) Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions:

```{r, results='hide'}
ggplot(  data = mpg,  mapping = aes(x = displ, y = hwy, color = drv) ) +  geom_point() +  geom_smooth(se = FALSE)
```
it creates a scatterplot and then adds a smooth line on top

## Question 3) What does show.legend = FALSE do? What happens if you remove it? Why do you think I used it earlier in ## the chapter? 

it remove the legend labels off the graph. Youu did not use it because it was unnessary. there were only 3 categories within that variable, and the color argument identify them already.

## Wuestion 4) What does the se argument to geom_smooth() do?

it display confidence interval around the smooth line

# EXERCISE 5

## Question 1) What is the default geom associated with stat_summary()? 
How could you rewrite the previous plot to use that geom function instead of the stat function?

geom_pointrange is the one. I would write it like this:

```{r,results='hide', error=TRUE}
ggplot(data = diamonds) + 
  geom_pointrange( mapping = aes(x = cut,y = depth),    
                  fun.ymin = min,    
                  fun.ymax =max,    
                  fun.y = median  )
```


## Question 2) What does geom_col() do? How is it different to geom_bar()? 

geom_col make the height of the bar with the values in the dataset, where as geom_bar make the height base on the proportion of the data values


## Question 3) Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?

They do the same thing

## Question 4) What variables does stat_smooth() compute? What parameters control its behavior? 

it computes the y variables. the mapping parameter, position and others things control its behaviors

Question 5) In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs? 

```{r, results='hide'}
ggplot(data = diamonds) +  geom_bar(mapping = aes(x = cut, y = ..prop..)) 
ggplot(data = diamonds) +  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..))
```

it's because proportion is a percentage and it cant go over 1

# EXERCISE 6

## Question 1) What is the problem with this plot? 
How could you improve it?

```{r, results='hide'}

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +  geom_point()

```

This plot has no main title, x-axis and y-axis does not tell you anything meanful about what the graph is about. The fix it I would just  add the things I mentioned.

## Question 2) What parameters to geom_jitter() control the amount of jittering? 

width and height


## Question 3) Compare and contrast geom_jitter() with geom_count().

jitter It adds a small amount of random variation to the location of each point, where as
geom_counts the number of observations at each location, then maps the count to point area



## Question 4) What’s the default position adjustment for geom_boxplot()? Create a visualization of the mpg dataset that demonstrates it.

# EXERCISE 7

## Question 1) Turn a stacked bar chart into a pie chart using coord_polar(). 
```{r}
bar <- ggplot(data = diamonds) +  geom_bar(    mapping = aes(x = cut, fill = cut),    show.legend = FALSE,    width = 1  ) +  theme(aspect.ratio = 1) +  labs(x = NULL, y = NULL)
bar + coord_polar()

```

## Questionn 2) What does labs() do? Read the documentation. 

They label your plots

## Question 3) What’s the difference between coord_quickmap() and coord_map()? 

coord_map projects a portion of the earth, which is approximately spherical, onto a flat 2D plane using any projection defined by the mapproj package. Map projections do not, in general, preserve straight lines, so this requires considerable computation. coord_quickmap is a quick approximation that does preserve straight lines. It works best for smaller areas closer to the equator.

## Question 4) What does the following plot tell you about the relationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do? 

there is a positive relationship. geom_abline() adds a trendline t the scaterplot. coord_fixed() is important because it adjust your cordinate system acording to your graph. 

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +  geom_point() +  geom_abline() +  coord_fixed()
```
